Probabilistic Forecasts in Hierarchical~Time~Series

### 2018-06-27 ProbabilisticCoherence.tex
Minor changes to section 5

Let $\bm{P}$ be a forecast distribution and let $\bm{Q}$ be the true data generating process respectively. Further let $\omega$ be a realization from $\bm{Q}$. Then a scoring rule is a function that maps $P,\omega$ to $\mathbb{R}$. It is a ``proper'' scoring rule if 
\begin{equation}\label{eq:(3.1.)}
\E_{\bm{Q}}[S(P,\omega)] \le \E_{\bm{Q}}[S(Q,\omega)],
\end{equation}
where $\E_{\bm{Q}}[S(\bm{P,\omega})]$ is the expected score under the true distribution $\bm{Q}$ \citep{Gneiting2008, Gneiting2014}.  When this inequality is strict, the scoring rule is said to be strictly proper.

In the context of probabilistic forecast reconciliation there could be two motivations for using scoring rules.  The first is to compare unreconciled densities to reconciled densities.  Although reconciliation is a valuable goal in and of itself since it can be important in aligning decision making across, for example, different units of an enterprise, in the point forecasting literature, forecast reconciliation has also been shown to improve forecast performance \todo{Some References}.  It will be worthwhile to see whether the same holds in the probabilistic forecasting case \todo{mention energy score proof - if we get it}.  The second motivation for using scoring rules is to compare two or more sets of reconciled probabilistic forecasts.  The objective here is to evaluate which reconciliation mapping $g(.)$ works best in practice.

@@

One way to evaluate probabilistic forecasts is via the application of univariate scoring rules to each variable in a hierarchy.  A summary can be taken of the expected scores across each margin for example a mean or median.  In the simulations of section~\ref{sec:gaussian} we consider two scoring rules.  The log score is given by the log of the marginal density of each variable.  The cumulative rank probability score generalises mean square error and is given by
\begin{align} \label{eq:CRPS}
\text{CRPS}(\breve{F}_i,y_{T+h,i}) &=\int \left(F_i(\breve{y}_i)-\mathbb{1}(y_i<y_{T+h,i})\right)dy_i\\ &=\E_{\breve{F}_i}|\breve{Y}_{T+h,i}-y_{T+h,i}| - \frac{1}{2}\E_{\breve{F}_i}|\breve{Y}_{T+h,i}-\breve{Y}^*_{T+h,i}|\,,
\end{align}
where the expectations in the second line can be approximated by Monte Carlo.

Although univariate scoring ruled have been used in the limited literature on probabilistic forecasting for hierarchies \cite{BenTaieb2017} and Jeon et al \todo{Jeon Panagiotelis and Petropoulos reference} to date, there are a number of shortcomings to this approach.  Crucially, evaluating univariate scores on the margins do not account for the dependence in the hierarchy.

@@

While the a number of alternative proper scoring rules are available for univariate forecasts, the multivariate case is somewhat more limited.  Here we focus on three cases, the log score, the energy score and the variogram score.  These are summarised in table~\ref{table:scoringrules} summarizes a few existing proper scoring rules.

@@

The log score can be approximated using a sample of values from the probabilistic forecast density \citep{Jordan2017} however it is more commonly used when a parametric form for the density is available for the probabilistic forecast.  So far, we have only defined probabilistic forecasts in terms of probability measures.  Although densities can be obtained for both reconciled and unreconciled forecasts, the degeneracy of reconciled forecasts is problematic when using log scores. We will discuss this further in the next subsection.

@@

Consider a rotated version of hierarchical time series $\bm{z}_t={\bm U}{\bm y}_t$ so that the first $m$ elements of $\bm{z}_t$ denoted $\bm{z}^(1)_t$ are unconstrained, while the remaining $n-m$ elements denoted $\bm{z}^(2)_t$ equal $0$ when the aggregation constraints hold. An example of the $n\times n$ ${\bm U}$ could be the left singular values of the matrix ${\bm S}$.  For a non-degenerate probability measure on $\mathbb{R}^n$, the density is the Radon-Nikodym derivative with respect to the usual Lesbegue measure on $\mathbb{R}^n$.  For a coherent probability measure the density after rotation is the Radon-Nikodym derivative with respect to the usual Lesbegue measure on $\mathbb{R}^m$.

Consider the case where the true density is $f_1(\bm{z}^{(1)}_t)\mathbb{1}\left({\bm{z}^{(2)}_t}\right)$ , the reconciled density is identical to the true density and the unreconciled density is given by $f_1(\bm{z}^{(1)}_t)f_2(\bm{z}^{(2)}_t)$, where $f_2$ is highly concentrated around $0$ but still a proper density, for example a Gaussian with variance $\sigma^2{\bm I}$ with $\sigma^2 < (2\pi)^{-1}$. The log score of the reconciled density is
$$
S\left(\tilde{f},\bm{z}^{(1)}_t\right) = -\log f_1\left(\bm{z}^{(1)}_t\right),
$$
while that of the unreconciled density is
\begin{align}
S\left(\hat{f},\bm{z}^{(1)}_t\right) &= -\log f_1(\bm{z}^{(1)}_t)-f_2(\bm{z}^{(1)}_t)\\
&= -\log f_1(\bm{z}^{(1)}_t)+\frac{n-m}{2}\log(2\pi\sigma^2)\\
&\leq-\log f_1(\bm{z}^{(1)}_t).
\end{align}

@@

For the log score, suppose the coherent probabilistic forecast has density $f(\bm{b})$.  The density for the full hierarchy is given by $f(\bm{y})=f(\bm{Sb})=f(\bm{b})J$ where $J=\prod_{j=1}^{m}\lambda_j^{-1}$ is the inverse of a pseudo-determinant of the non-square  matrix $X$ and $\lambda_j$ are the non-zero singular values of ${\bm S}$.  Therefore for any coherent density the log score of the full hierarchy differs from the log score for the bottom level series by the term $log(J)$.  This term depends  only on the structure of the  hierarchy and is fixed across different forecasting methods.  Therefore if one method achieves a lower expected log score than an alternative method  when assessed using the bottom level series, the same ordering is preserved when an assessment is made on the basis of the full hierarchy.

The same property does not hold for all scores in general.  For example, energy score can be expressed in terms of an expectations of norms.  In general, since norms are invariant under orthogonal rotations the energy score is also invariant under orthogonal transformations \citep{Szekely2013,Gneiting2007}.  In the context of two coherent forecasts the same is true of a semi-orthogonal transformation from a lower dimensional basis series to the full hierarchy.  However,  when ${\bm S}$ is  the usual summing matrix, it is not semi-orthogonal.  As such the energy score computed on the bottom level series will differ from the energy score computed using the full hierarchy and the ordering of different reconciliation methods may change depending on the basis series used.  In this case we recommend computing energy score using the full hierarchy.  Although the discussion here is related to energy score, the same logic holds for other scores that are not invariant to orthogonal rotations, for example the variogram score.

The properties of mutlivariate scoring rules in the context of evaluating reconciled probabilistic forecasts in Table~\ref{tab:prop}.

>>>>

Let $P$ be a forecast distribution and let $Q$ be the true data generating process respectively. Furthermore let $\omega$ be a realization from $Q$. Then a scoring rule is a function $S(P,\omega)$ that maps $P,\omega$ to $\mathbb{R}$. It is a ``proper'' scoring rule if 
\begin{equation}\label{eq:(3.1.)}
\E_{\bm{Q}}[S(P,\omega)] \le \E_{\bm{Q}}[S(Q,\omega)],
\end{equation}
where $\E_{Q}[S(P,\omega)]$ is the expected score under the true distribution $Q$ \citep{Gneiting2008, Gneiting2014}.  When this inequality is strict, the scoring rule is said to be strictly proper.

In the context of probabilistic forecast reconciliation there could be two motivations for using scoring rules.  The first is to compare unreconciled densities to reconciled densities.  Although reconciliation is a valuable goal in and of itself since it can be important in aligning decision making across, for example, different units of an enterprise, in the point forecasting literature, forecast reconciliation has also been shown to improve forecast performance \todo{Some References}.  It will be worthwhile to see whether the same holds in the probabilistic forecasting case.  The second motivation for using scoring rules is to compare two or more sets of reconciled probabilistic forecasts to one another.  The objective here is to evaluate which reconciliation mapping $g(.)$ works best in practice.

@@

One way to evaluate probabilistic forecasts is via the application of univariate scoring rules to each variable in a hierarchy.  A summary can be taken of the expected scores across each margin for example a mean or median.  In the simulations of section~\ref{sec:gaussian} we consider two scoring rules.  The log score is given by the log of the marginal density of each variable.  The cumulative rank probability score generalises mean square error and is given by
\begin{align} \label{eq:CRPS}
\text{CRPS}(\breve{F}_i,y_{T+h,i}) &=\int \left(\breve{F}_i(\breve{y}_i)-\mathbb{1}(y_i<y_{T+h,i})\right)dy_i\\ &=\E_{\breve{Y}_i}|\breve{Y}_{T+h,i}-y_{T+h,i}| - \frac{1}{2}\E_{\breve{Y}_i}|\breve{Y}_{T+h,i}-\breve{Y}^*_{T+h,i}|\,,
\end{align}
where $\breve{Y}_i$ and $\breve{Y}^*_{T+h,i}$ are independent copies of a random variable with distribution $\breve{F}_i$ and the latter is the predictive distribution for the $i^{th}$ margin of a forecast for time $t+h$ made at time $t$. The expectations in the second line can be approximated by Monte Carlo when a sample from the predictive distribution is available.

An advantage to this approach is that it allows the forecaster to evaluate the levels and individual series of the hierarchy where the gains to reconciliation are greatest.  For this reason this approach has been used in the limited literature on probabilistic forecasting for hierarchies \cite{BenTaieb2017} and Jeon et al \todo{Jeon Panagiotelis and Petropoulos reference} to date.  A major shortcomings to this approach however is that, evaluating univariate scores on the margins do not account for the dependence in the hierarchy.

@@

While the a number of alternative proper scoring rules are available for univariate forecasts, the multivariate case is somewhat more limited.  Here we focus on three scoring rules: the log score, the energy score and the variogram score.  These are summarized in table~\ref{table:scoringrules}.

@@

The log score can be approximated using a sample of values from the probabilistic forecast density \citep{Jordan2017} however it is more commonly used when a parametric form for the density is available for the probabilistic forecast.  So far, we have mainly defined probabilistic forecasts in terms of probability measures.  Although densities can be obtained for both reconciled and unreconciled forecasts, the degeneracy of reconciled forecasts is problematic when using log scores. We will discuss this further in the next subsection.

@@

Consider a rotated version of hierarchical time series $\bm{z}_t={\bm U}{\bm y}_t$ so that the first $m$ elements of $\bm{z}_t$ denoted $\bm{z}^{(1)}_t$ are unconstrained, while the remaining $n-m$ elements denoted $\bm{z}^{(2)}_t$ equal $0$ when the aggregation constraints hold. An example of the $n\times n$ ${\bm U}$ could be the matrix of left singular vectors of ${\bm S}$.  For a non-degenerate probability measure on $\mathbb{R}^n$, the density is the Radon-Nikodym derivative with respect to the usual Lesbegue measure on $\mathbb{R}^n$.  %For a coherent probability measure the density after rotation is the Radon-Nikodym derivative with respect to the usual Lesbegue measure on $\mathbb{R}^m$.

Consider the case where the true density is $f_1(\bm{z}^{(1)}_t)\mathbb{1}\left({\bm{z}^{(2)}_t}\right)$, and is compared to an incoherent density is given by $f_1(\bm{z}^{(1)}_t)f_2(\bm{z}^{(2)}_t)$, where $f_2$ is highly concentrated around $0$ but still a proper density. For example $f_2$ may be Gaussian with variance $\sigma^2{\bm I}$ with $\sigma^2 < (2\pi)^{-1}$. The log score under the true DGP is
$$
S\left(\tilde{f},\bm{z}^{(1)}_t\right) = -\log f_1\left(\bm{z}^{(1)}_t\right),
$$
while that of the unreconciled density is
\begin{align}
S\left(\hat{f},\bm{z}^{(1)}_t\right) &= -\log f_1(\bm{z}^{(1)}_t)-f_2(\bm{z}^{(1)}_t)\\
&= -\log f_1(\bm{z}^{(1)}_t)+\frac{n-m}{2}\log(2\pi\sigma^2)\\
&<-\log f_1(\bm{z}^{(1)}_t).
\end{align}

@@

For the log score, suppose the coherent probabilistic forecast has density $f(\bm{b})$.  The density for the full hierarchy is given by $f(\bm{y})=f(\bm{Sb})=f(\bm{b})J^{-1}$ where $J=\prod_{j=1}^{m}\lambda_j$ is a pseudo-determinant of the non-square  matrix ${\bm S}$ and $\lambda_j$ are the non-zero singular values of ${\bm S}$.  Therefore for any coherent density the log score of the full hierarchy differs from the log score for the bottom level series by the term $log(J)$.  This term depends  only on the structure of the  hierarchy and is fixed across different reconciliation methods.  Therefore if one method achieves a lower expected log score compared to an alternative method  when assessed using the bottom level series, the same ordering is preserved when an assessment is made on the basis of the full hierarchy.

The same property does not hold for all scores in general.  For example, energy score can be expressed in terms of expectations of norms.  In general, since norms are invariant under orthogonal rotations the energy score is also invariant under orthogonal transformations \citep{Szekely2013,Gneiting2007}.  In the context of two coherent forecasts the same is true of a semi-orthogonal transformation from a lower dimensional basis series to the full hierarchy.  However,  when ${\bm S}$ is  the usual summing matrix, it is not semi-orthogonal.  As such the energy score computed on the bottom level series will differ from the energy score computed using the full hierarchy and the ordering of different reconciliation methods may change depending on the basis series used.  In this case we recommend computing energy score using the full hierarchy.  Although the discussion here is related to energy score, the same logic holds for other multivariate scores that are not invariant to orthogonal rotations, for example the variogram score.

The properties of multivariate scoring rules in the context of evaluating reconciled probabilistic forecasts in Table~\ref{tab:prop}.
